Design and Implementation of Probabilistic Programming Language Anglican

### 2016-07-07 paper/paper.tex
language definition

\section{Anglican Syntax}

\cite{SDF+10, PC94, H08}

@@

Literals are the same in Anglican and Clojure. They are 
left unmodified. Literals are a subset of opaque expressions,
and are identified by test \texttt{simple?} called from
\texttt{opaque?}.  However, The Clojure syntax has a peculiarity
of using the syntax of compound literals (vectors, hash maps,
and sets) for data constructors. Hence, compound literals must
be traversed recursively, and if there is a nested non-literal
component, transformed into a call to the corresponding data
constructor. Functions \texttt{cps-of-vector},
\texttt{cps-of-hash-map}, \texttt{cps-of-set}, called from
\texttt{cps-of-expression}, transform Clojure constructor syntax
(\texttt{[...]}, \texttt{\{...\}}, \texttt{\#\{...\}}) into the
corresponding calls.

>>>>

\section{Language}

The Anglican language is a subset of Clojure\footnote{It would be
possible to support almost full Clojure by expanding all macros
in the Anglican source code, however in Clojure, unlike in
Scheme~\cite{SDF+10}, or Common Lisp~\cite{PC94}, the result
of macro-expansion of derived special forms is not well
specified and implementation specific.}.  Within \texttt{defquery},
\texttt{let}, \texttt{if}, \texttt{when}, \texttt{cond},
\texttt{case}, \texttt{and}, \texttt{or}, \texttt{fn} forms are
supported (others may be in the future but are not now). 
In \texttt{let} bindings and \texttt{fn} argument lists,
vector destructuring (but not hash map destructuring) is
supported. Compound literals for vectors, hash maps, and
sets are supported just like in Clojure.

\subsection{\texttt{recur}}

Anglican is stackless, therefore \texttt{recur} is
unnecessary, no recursive call can lead to stack overflow;
Recursive calls to functions should be used instead. However,
\texttt{loop}/\texttt{recur} is provided for convenience as a way to express
loops. \texttt{recur} outside of \texttt{loop} will lead to unpredictable
behaviour and hard-to-catch errors.

\subsection{Core library}

All of Clojure's core library except for higher-order functions
(functions that accept other functions as arguments) is
available in Anglican. Higher order functions cannot be reused
from Clojure, they have to be re-implemented to accept
functional arguments in CPS form. The following higher-order
functions are implemented: \texttt{map}, \texttt{reduce},
\texttt{filter}, \texttt{some}, \texttt{repeatedly},
\texttt{comp}, \texttt{partial}.

@@

Literals are the same in Anglican and Clojure. They are 
left unmodified. Literals are a subset of opaque expressions,
and are identified by test \texttt{simple?} called from
\texttt{opaque?}.  However, the Clojure syntax has a peculiarity
of using the syntax of compound literals (vectors, hash maps,
and sets) for data constructors. Hence, compound literals must
be traversed recursively, and if there is a nested non-literal
component, transformed into a call to the corresponding data
constructor. Functions \texttt{cps-of-vector},
\texttt{cps-of-hash-map}, \texttt{cps-of-set}, called from
\texttt{cps-of-expression}, transform Clojure constructor syntax
(\texttt{[...]}, \texttt{\{...\}}, \texttt{\#\{...\}}) into the
corresponding calls.
